---
title: "Various Jail Population"
output: html_document
date: '2022-05-20'
---

## Introduction 

From the incarceration dataset, numerous domains need to be expressed as with this assignment. The focus will be on the issue of the total jail population in different divisions, regions, and states. With the problem domain, we will review the most crucial factor of race as we will look into black, white, and other races and how each race contributes to the total jail population. As we gain information on these values through charts and maps, it allows us to gain a general view of what is happening in society and why the issue of racism and inequality is expressed all the time and has not changed throughout the years. Overall, with the assignment, it is essential to view all this information as we can learn more about how to make a better society for future generations and prevent the same results each year. 

## Summary Information 

```{r echo=FALSE, message=FALSE}
source("a3 analysis.r")
summary_info
```

#### Summary Paragraph: 

The data has numerous variables as it contains 121 different variables that they collected. With the summary information, it embraces the dataset's fundamental areas they are collecting from, such as the division, state, and region, as it exposes the area that plays a factor in the jail population. The combination of each place calculated the total average of 161.1213 jail population, and in 2018 the total max jail population was 17208. With these initial, I wanted to know how the primary two races play a role in the total population, and as we can tell, the max black jail population was 13154.92, and the max white jail population was 7036.59 where it has a difference of 6107.33. Without a surprise, we can realize that racial difference varies based on race, as those two primaries significantly impact the jail population more than other races. The max jail population is just eight, whereas people outside these two races do not experience jail as much. These values highlight the terms of inequality and the social norm of society. 

## Bar Chart on Trends Over Time

```{r echo=FALSE, warning = FALSE, message = FALSE}
source("a3 analysis.r")
plot(barchart)
```

#### Observations:

The bar chart was chosen to represent the jail population throughout the years in urban cities. The chart emphasizes a trend over time in the four urban cities: rural cities, small to mid-cities, suburban cities, and urban cities captured through the dataset. We can view how different cities make a difference in the jail population with the chart. The chart shows that more populated cities have the most jail population compared to smaller cities with fewer individuals. Also, the chart emphasizes a pattern of the four urban cities being constant, with the jail population as not one area has a dramatic spike. However, we can see that urban cities' jail population has decreased throughout the years. Finally, it gives the essential purpose of allowing people to realize that living in a smaller populated area is safer. 

## Scatter Plot Comparing Two Variables 

```{r, echo=FALSE}
source("a3 analysis.r")
plot(scatterplot)
```

#### Observations:

The scatter plot's purpose emphasized the racial difference between the black-to-white jail population in the different regions. The chart enables viewers to recognize the inequality of the racial difference as it shows that the black community jail population is significantly more than the white community jail population throughout each region. Wherein the northeast expresses a dramatic increase in black than white as mainly only black experience jail time. The display embraces the concept of blacks being more violent than white and continues the narrative that black people are more aggressive with their actions. It contributes to our social norm that black people are dangerous, causing the mistreatment to continue and the unfairness they experience compared to white people. 

## Map 

```{r, echo=FALSE}
source("a3 analysis.r")
plot(map)
```

#### Observations:

The map imposes the purpose of displaying the other race contributing to the calculation of the total jail population. The addition of the map helps identify which state has the most impact on the jail population. In contrast, it shows that most states either had no calculation due to the data collector not collecting from that state which is shown as grey, and the rest are mainly green where the amount of jail population is meager for people of other races. However, California and Minnesota had a darker green than the rest as these two areas are more aggressive. It results in more people ending in jail, adding to the jail population, including other races. This is essential for including the map and what it displays because these races usually do not get recognized as the other two primary races of black and white get most of the attention due to circumstances like these wherein discussions of other races do not get included, such as equality where other races do get mistreated as well. 


